---
title: "EDS 213 Week 2 Assignment 2.1 & 2.2"
author: "Halina Do-Linh"
date: "10/6/2021"
output: 
  html_document:
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(dataRetrieval)
library(metajam)
library(here)
```

# Assignment 2.1 - `dataRetrieval`
```{r}
siteNumber <- c("11118500", "11113500") # added santa paula's site number 
parameterCd <- "00060" # discharge number is standard
startDate <- "2019-10-01"
endDate <- "2020-10-05"

vta_sp_discharge <- readNWISdv(siteNumber, parameterCd, startDate, endDate) %>% 
  dplyr::rename("Discharge" = X_00060_00003)

vta_sp_discharge_plot <- ggplot(data = vta_sp_discharge, aes(x = Date, y = Discharge, color = site_no)) +
  geom_line() +
  scale_color_manual(name = "Site Id", labels = c("Santa Paula", "Ventura"), values = c("red", "cyan4"))

vta_sp_discharge_plot
```

**What is the best way to query multiple sites using this R package?** 

- By creating a vector with the site numbers that you want, assigning this vector as an object, and then calling that object in the `readNWISdv()` function. 

# Assignment 2.2 - `metajam`
```{r, message=FALSE}
# set inputs
data_url <- "https://knb.ecoinformatics.org/knb/d1/mn/v2/object/urn%3Auuid%3A7fc6f6db-c5ea-426a-a743-1f2edafb43b8"

# metajam::download_d1_data(data_url, here::here("data")) 
# had to comment out metajam::download_d1_data() in order to knit 

alaska_metadata <- read_d1_files(here::here("data", "Alaska-HouseholdLanguage-csv"))

hh_data <- alaska_metadata$data # view data frame from alaska_metadata list
```

**Write a piece of code that will compute the percentage of Alaskan household speaking only English for the year 2009 to 2015**
```{r, message=FALSE}
eng_lang <- hh_data %>% 
  select("Year", "city", "SASAP.Region", "lng", "lat", "total", "speak_only_english") %>%
  group_by(Year) %>% 
  summarize(pct_eng_lang = sum(speak_only_english) / sum(total))

average_2009_2015 <- mean(eng_lang$pct_eng_lang)
```

The percent average of Alaskan households speaking only English from 2009 to 2015 is approximately 77%. 
